---
title: "SafiyaFinalProjectScript"
author: "Safiya Alavi"
date: "10/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(tidyverse)
library(tidymodels)
library(corrplot)
library(ggthemes)
library(klaR) # for naive bayes
library(discrim)
library(glmnet)
library(rpart)
library(randomForest)
library(xgboost)
library(ranger)
library(vip)
library(lubridate)
library(dplyr)
library(ISLR)
library(rpart.plot)
library(janitor)
library(ggpubr)

tidymodels_prefer()

# LDA 
load("WhiteLDA.rda")
load("RedLDA.rda")

# LDA with PCA
load("WhiteLDAPCA.rda")
load("RedLDAPCA.rda")

# Decision Tree
load("WhiteWineDecisionTree.rda")
load("RedWineDecisionTree.rda")

# Random Forest 
load("WhiteWineRandomForest.rda")
load("RedWineRandomForest.rda")

# Boosted Trees
load("WhiteWineBoostedTrees.rda")
load("RedWineBoostedTrees.rda")
```

## Overview of Data Set

<br>

We are analyzing the quality of wine based on 11 predictors: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulfates, alcohol. Quality, in this instance, is defined as an indicator of its craftsmanship and, thus, desirability which can be used, for example, in pricing. Quality does not necessarily indicate if a wine has gone bad. We will use two data sets -- one for white wine and the other for red -- to create two respective machine learning models. There are 4899 observations in the white wine data set and 1600 observations in the red wine data set.

## Overview of Research Questions

<br>

We are interested in conducting analysis on both the white and red wine to assess the quality of wine in new wines in production so that they may be accurately priced and marketed.

## Proposed Project Timeline

<br>

Week 2: Data cleaning<br> Week 3-5: Exploratory Data Analysis<br> Week 6-8: Model Building<br> Week 9: Conclusion and final touches<br>

## Any questions or concerns

<br>

# Loading Data and Packages

This project uses data on the white and red wine, which records information about the details of the wine.

```{r load data, message = FALSE}

white <- read.csv("Wine Dataset/winequality-white.csv", sep = ";")

red <- read.csv("Wine Dataset/winequality-red.csv", sep = ";")

# Make Column Names Clean 
#colnames(white) <- c("Fixed_Acidity", "Volatile_Acidity","Citric_Acid", "Residual_Sugar", "Chlorides", "Free_Sulfur_Dioxide", "Total_Sulfur_Dioxide", "Density", "pH", "Sulfates", "Alcohol", "Quality")
#colnames(red) <- c("Fixed_Acidity", "Volatile_Acidity","Citric_Acid", "Residual_Sugar", "Chlorides", "Free_Sulfur_Dioxide", "Total_Sulfur_Dioxide", "Density", "pH", "Sulfates", "Alcohol", "Quality")
white %>% clean_names()
red %>% clean_names()

white <- white[white$quality < 9,]


# Change quality to factor
white$quality <- factor(white$quality, levels = c(3,4,5,6,7,8))
red$quality <- factor(red$quality, levels = c(3,4,5,6,7,8))

# Adding type
white$type <- "White"
red$type <- "Red"

# data frame of combined wine
combinedWine <- rbind(white, red)
```

# Data Cleaning

Our data does not need to be cleaned because it's perfect :D

# Data Split

```{r data split, message = FALSE}
set.seed(1234)
white_split <- white %>% 
  initial_split(prop = 0.7, strata = "quality")

white_train <- training(white_split)
white_test <- testing(white_split)

red_split <- red %>% 
  initial_split(prop = 0.7, strata = "quality")

red_train <- training(red_split)
red_test <- testing(red_split)

white_fold <- vfold_cv(white_train, v = 5)
red_fold <- vfold_cv(red_train, v = 5)
```

## Exploratory Data Analysis

```{r eda correlation matrix, message = FALSE}
white %>% 
  select(is.numeric) %>% 
  cor() %>% 
  corrplot(type = 'lower', diag = FALSE, 
           method = 'color', main = 'White Wine Correlation Plot')

red %>% 
  select(is.numeric) %>% 
  cor() %>% 
  corrplot(type = 'lower', diag = FALSE, 
           method = 'color', main = 'Red Wine Correlation Plot')
```

```{r eda boxplot of alcohol versus quality, message = FALSE}
ggplot(white_train, aes(x = alcohol, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "White Wine Alcohol Content versus Quality", fill = "Quality")  +
    scale_fill_brewer(palette="BuPu")

ggplot(red_train, aes(x = alcohol, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "Red Wine Alcohol Content versus Quality", fill = "Quality")  +
    scale_fill_brewer(palette="RdPu")

ggplot(combinedWine, mapping = aes(x = alcohol, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Wine Alcohol Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd") + facet_wrap(. ~ type, nrow = 1)

ggplot(combinedWine, mapping = aes(x = alcohol, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Wine Alcohol Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="OrRd") 
```

```{r eda boxplot of fixed acidity, volatile acidity, and citric acid versus quality, message = FALSE}

# FIXED ACIDITY 
# White
ggplot(white_train, aes(x = fixed.acidity, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "White Wine Fixed Acidity versus Quality", fill = "Quality")  +
    scale_fill_brewer(palette="BuPu") + coord_cartesian( xlim = c(0,16), ylim = NULL, default = FALSE )

# Red
ggplot(red_train, aes(x = fixed.acidity, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "Red Wine Fixed Acidity versus Quality", fill = "Quality")  +
    scale_fill_brewer(palette="RdPu") + coord_cartesian( xlim = c(0,16), ylim = NULL, default = FALSE )

# Same Page but Seperate
ggplot(combinedWine, mapping = aes(x = `fixed.acidity`, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Fixed Acidity Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd") + facet_wrap(. ~ type, nrow = 1) +coord_cartesian( xlim = c(0,16), ylim = NULL, default = FALSE )

# Combined
ggplot(combinedWine, mapping = aes(x = `fixed.acidity`, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Fixed Acidity Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd")+coord_cartesian( xlim = c(0,16), ylim = NULL, default = FALSE )
# ___________________________________________________________________________
# VOLATILE ACIDITY 
# White
ggplot(white_train, aes(x = `volatile.acidity`, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "White Wine Volatile Acidity versus Quality", fill = "Quality")  + scale_fill_brewer(palette="BuPu")+ coord_cartesian( xlim = c(0,1), ylim = NULL, default = FALSE )

# Red
ggplot(red_train, aes(x = `volatile.acidity`, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "Red Wine Volatile_Acidity versus Quality", fill = "Quality")  +
    scale_fill_brewer(palette="RdPu") + coord_cartesian( xlim = c(0,1), ylim = NULL, default = FALSE )

# Same Page but Seperate
ggplot(combinedWine, mapping = aes(x = `volatile.acidity`, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Volatile_Acidity Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd") + facet_wrap(. ~ type, nrow = 1) + coord_cartesian( xlim = c(0,1), ylim = NULL, default = FALSE )

# Combined
ggplot(combinedWine, mapping = aes(x = `volatile.acidity`, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Volatile_Acidity Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd") +coord_cartesian( xlim = c(0,1), ylim = NULL, default = FALSE )

# ___________________________________________________________________________
# CITRIC ACID 
# White
ggplot(white_train, aes(x = `citric.acid`, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "White Wine Citric Acid Levels versus Quality", fill = "Quality")  + scale_fill_brewer(palette="BuPu")

# Red
ggplot(red_train, aes(x = `citric.acid`, y = quality)) + geom_boxplot(aes(fill = quality))+ geom_point() + 
  labs(title = "Red Wine Citric Acid Levels versus Quality", fill = "Quality")  + scale_fill_brewer(palette="RdPu")

# Same Page but Separate (Cut off at 1)
ggplot(combinedWine, mapping = aes(x = `citric.acid`, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Citric Acid Levels Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd") + facet_wrap(. ~ type, nrow = 1)+ coord_cartesian( xlim = c(0,1), ylim = NULL, default = FALSE )

# Combined
ggplot(combinedWine, mapping = aes(x = `citric.acid`, y = quality, fill = quality)) + geom_boxplot() + geom_point(white_train, mapping = aes(x = alcohol, y = quality)) + labs(title = "Red and White Citric Acid Levels Content versus Quality", fill = "Quality")  + scale_fill_brewer(palette="PuRd")+ coord_cartesian( xlim = c(0,1), ylim = NULL, default = FALSE )
```

```{r eda scatter plots}
# scaled datasets 

scaled_white = as.data.frame(scale(select(white, c(-quality,-type))))
scaled_red = as.data.frame(scale(select(red, c(-quality,-type))))

# WHITE
# scaled white residual sugar versus density 
wrsd = ggplot(scaled_white, aes(x = residual.sugar, y = density)) + geom_point()+scale_x_continuous(name = "Residual Sugar") + scale_y_continuous(name = "Density") + geom_smooth(method = "lm", se = FALSE)+ ggtitle(" Residual Sugar Versus Density") + theme(plot.title = element_text(size = 5))
# scaled white alcohol versus density 
wad = ggplot(scaled_white, aes(x = alcohol, y = density)) + geom_point()+scale_x_continuous(name = "Alcohol") + scale_y_continuous(name = "Density") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Alcohol Versus Density") + theme(plot.title = element_text(size = 5))

# scaled white free sulfur dioxide versus total sulfur dioxide
wfsdtsd = ggplot(scaled_white, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) + geom_point()+scale_x_continuous(name = "Free Sulfur Dioxide") + scale_y_continuous(name = "Total Sulfur Dioxide") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Free Sulfur Versus Total Sulfur Dioxide") + theme(plot.title = element_text(size = 5))

ggarrange(wrsd, wad, wfsdtsd + rremove("x.text"), ncol = 3, nrow = 1) %>% annotate_figure(top = text_grob("White Wine Scaled"))
# RED

# scaled red volatile acidity versus citric acid
rvaca = ggplot(scaled_red, aes(x = volatile.acidity, y = citric.acid)) + geom_point()+scale_x_continuous(name = "Volatile Acidity") + scale_y_continuous(name = "Citric Acid") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Volatile Acidity Versus Citric Acid") + theme(plot.title = element_text(size = 5))

# scaled red fixed acidity versus citric acid 
rfaca = ggplot(scaled_red, aes(x = fixed.acidity, y = citric.acid)) + geom_point()+scale_x_continuous(name = "Fixed Acidity") + scale_y_continuous(name = "Citric Acid") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Fixed Acidity Versus Citric Acid") + theme(plot.title = element_text(size = 5))

# scaled red fixed acidity versus pH
rfap = ggplot(scaled_red, aes(x = fixed.acidity, y = pH)) + geom_point()+scale_x_continuous(name = "Fixed Acidity") + scale_y_continuous(name = "pH") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Fixed Acidity Versus pH") + theme(plot.title = element_text(size = 5))

# scaled red fixed acidity versus density
rfad = ggplot(scaled_red, aes(x = fixed.acidity, y = density)) + geom_point()+scale_x_continuous(name = "Fixed Acidity") + scale_y_continuous(name = "Density") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Fixed Acidity Versus Density") + theme(plot.title = element_text(size = 5))

# scaled red free sulfur dioxide versus total sulfur dioxide
fsdtsd = ggplot(scaled_red, aes(x = free.sulfur.dioxide, y = total.sulfur.dioxide)) + geom_point()+scale_x_continuous(name = "Free Sulfur Dioxide") + scale_y_continuous(name = "Total Sulfur Dioxide") + geom_smooth(method = "lm", se = FALSE)+ ggtitle("Free Sulfur Dioxide Versus Total Sulfur Dioxide") + theme(plot.title = element_text(size = 5))

ggarrange(rvaca, rfaca, rfap, rfad, fsdtsd, ncol = 3, nrow = 2) %>% annotate_figure(top = text_grob("Red Wine Scaled"))

```

```{r residual sugar red}
# we removed outliers to ensure that the variation was not due to the outliers 
ggplot(red[red$free.sulfur.dioxide < 50,], aes(x = free.sulfur.dioxide, y = quality)) + 
  geom_boxplot(aes(fill = quality)) +
  labs(title = "Free Sulfur Dioxide for Red Wine", x = "Free Sulfur Dioxide", y = "Quality") +
  geom_point(width = 0.15) +
  scale_fill_brewer(palette = "RdPu")

```

## Model Fitting

```{r creating recipe}
# creating recipes

white_recipe <- recipe(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = white_train) %>% step_dummy(all_nominal_predictors()) %>% step_normalize(all_predictors()) #%>% step_interact(~density:residual.sugar + density:alcohol) 

red_recipe <- recipe(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = red_train) %>% step_dummy(all_nominal_predictors()) %>% step_normalize(all_predictors())
#%>% step_interact(~ citric.acid:volatile.acidity + free.sulfur.dioxide:total.sulfur.dioxide)

```

```{r White Log, LDA, QDA, NB k-fold}
# white 

#lda model using cross validation
wlda_model <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS") 

wlda_wkflow<- workflow() %>% 
  add_model(wlda_model) %>% 
  add_recipe(white_recipe)

wlda_fit_cross <- fit_resamples(wlda_wkflow, white_fold)

collect_metrics(wlda_fit_cross)%>% mutate(model_type = "LDA")

#naive bayes model using cross validation
wnb_mod <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("klaR") %>% 
  set_args(usekernel = FALSE) 

wnb_wkflow <- workflow() %>% 
  add_model(wnb_mod) %>% 
  add_recipe(white_recipe)

wnb_fit_cross <- fit_resamples(wnb_wkflow, white_fold)

collect_metrics(wnb_fit_cross) %>% mutate(model_type = "Naive Bayes")
```

```{r Red Log, LDA, QDA, NB k-fold}
# red 

#lda model using cross validation
rlda_model <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS") 

rlda_wkflow<- workflow() %>% 
  add_model(rlda_model) %>% 
  add_recipe(red_recipe)

rlda_fit_cross <- fit_resamples(rlda_wkflow, red_fold)

collect_metrics(rlda_fit_cross) %>% mutate(model_type = "LDA")

#naive bayes model using cross validation
rnb_mod <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("klaR") %>% 
  set_args(usekernel = FALSE) 

rnb_wkflow <- workflow() %>% 
  add_model(rnb_mod) %>% 
  add_recipe(red_recipe)

rnb_fit_cross <- fit_resamples(rnb_wkflow, red_fold)

collect_metrics(rnb_fit_cross) %>% mutate(model_type = "Naive Bayes")
```

```{r log reg}
glm.fit <- glm(quality ~ volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = red_train, family = binomial)
summary(glm.fit)

# white wine
wlog_model <- logistic_reg() %>% 
  set_mode("classification")%>% 
  set_engine("glm")

wlog_wkflow <- workflow() %>% 
  add_model(wlog_model) %>% 
  add_recipe(white_recipe)

wlog_reg_fit <- fit(wlog_wkflow, white_train)

wlog_reg_res <- predict(wlog_reg_fit, new_data = white_test %>%  select(-quality), type = "class" ) 

wlog_reg_res <- bind_cols(wlog_reg_res, white_test %>%  select(quality)) 

wlog_reg_acc <- augment(wlog_reg_fit, new_data = white_test) %>% accuracy(truth = quality, estimate = .pred_class)

print(wlog_reg_res)
print(wlog_reg_acc)

# red wine
rlog_model <- logistic_reg() %>% 
  set_engine("glm") %>% set_mode("classification")

rlog_wkflow <- workflow() %>% 
  add_model(rlog_model) %>% 
  add_recipe(red_recipe)

rlog_reg_fit <- fit(rlog_wkflow, red_train)

rlog_reg_res <- predict(rlog_reg_fit, new_data = red_test %>%  select(-quality), type = "class" ) 

rlog_reg_res <- bind_cols(rlog_reg_res, red_test %>%  select(quality)) 

rlog_reg_acc <- augment(rlog_reg_fit, new_data = red_test) %>% accuracy(truth = quality, estimate = .pred_class)

print(rlog_reg_res)
print(rlog_reg_acc)

#collect_metrics(wlog_reg_fit)
#collect_metrics(rlog_reg_fit)
```

```{r LDA}

# white lda model
wlda_model <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS") 

# setting up workflow
wlda_wkflow <- workflow() %>% 
  add_model(wlda_model) %>% 
  add_recipe(white_recipe)

# fitting the model to training data
wlda_fit <- fit(wlda_wkflow, white_train)

# now testing the model on the testing data 
wlda_res <- predict(wlda_fit, new_data = white_test %>%  select(-quality), type = "class" ) 

wlda_res <- bind_cols(wlda_res, white_test %>%  select(quality)) 

# returning the roc auc, roc auc curves, heatmap 
wlda_pred <- augment(wlda_fit, new_data = white_test) 
wlda_rocauc <- wlda_pred %>% roc_auc(truth = quality, estimate = .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% mutate(model_type = "White LDA Model")
wlda_roccurve <- wlda_pred %>% roc_curve(quality, .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% autoplot()
wlda_confusionmatrix <- wlda_pred %>% conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

print(wlda_res)
print(wlda_rocauc)
wlda_roccurve
wlda_confusionmatrix

# red lda model
rlda_model <- discrim_linear() %>% 
  set_mode("classification") %>% 
  set_engine("MASS") 

rlda_wkflow <- workflow() %>% 
  add_model(rlda_model) %>% 
  add_recipe(red_recipe)

rlda_fit <- fit(rlda_wkflow, red_train)

rlda_res <- predict(rlda_fit, new_data = red_test %>%  select(-quality), type = "class" ) 

rlda_res <- bind_cols(rlda_res, red_test %>%  select(quality)) 

# returning the roc auc
rlda_pred <- augment(rlda_fit, new_data = red_test)
rlda_rocauc <- rlda_pred %>% roc_auc(truth = quality, estimate = .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8)%>% mutate(model_type = "Red LDA Model")
rlda_roccurve <- rlda_pred %>% roc_curve(quality, .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% autoplot()
rlda_confusionmatrix <- rlda_pred %>% conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

print(rlda_res)
print(rlda_rocauc)
rlda_roccurve
rlda_confusionmatrix
```

```{r Naive Bayes model}
# white
wnb_mod <- naive_Bayes() %>% 
  set_mode("classification") %>% 
  set_engine("klaR") %>% 
  set_args(usekernel = FALSE) 

wnb_wkflow <- workflow() %>% 
  add_model(wnb_mod) %>% 
  add_recipe(white_recipe)

wnb_fit_cross <- fit_resamples(wnb_wkflow, white_fold)

collect_metrics(wnb_fit_cross)
```

```{r decision tree white}
# decision tree specification
wtree_spec <- decision_tree() %>%
  set_engine("rpart", model=TRUE)

# setting mode to classification
wtree_spec_class <- wtree_spec %>%
  set_mode("classification")

wclass_tree_fit <- wtree_spec_class %>%
  fit(quality ~ volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = white_train)


wDecisionTreePre <- wclass_tree_fit %>%
  extract_fit_engine() %>%
  rpart.plot(main="White Wine Decision Tree")

# augmented on training 
wDecisionTreeAccPre <- augment(wclass_tree_fit, new_data = white_train) %>%
  accuracy(truth = quality, estimate = .pred_class)

wDecisionTreeConfMatrixPre <- augment(wclass_tree_fit, new_data = white_train) %>%
  conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

# augmentsed on testing 
augment(wclass_tree_fit, new_data = white_test) %>%
  conf_mat(truth = quality, estimate = .pred_class)  %>% autoplot(type = "heatmap")

augment(wclass_tree_fit, new_data = white_test) %>%
  accuracy(truth = quality, estimate = .pred_class)

# tuning cost complexity 
wclass_tree_wf<- workflow() %>%
  add_model(wtree_spec_class %>% 
              set_args(cost_complexity = tune())) %>% 
  add_formula(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol)

param_grid <- grid_regular(cost_complexity(range = c(-3,-1)), levels = 10)

tune_res_white <- tune_grid(
  wclass_tree_wf,
  resamples = white_fold,
  grid = param_grid,
  metric = metric_set(accuracy)
)

wAutoPlot <- autoplot(tune_res_white)

# extracting the best cost complexitiy parameter
wbest_rocauc <- collect_metrics(tune_res_white) %>% arrange(desc(mean))
best_complexity <- select_best(tune_res_white)

wclass_tree_final <- finalize_workflow(wclass_tree_wf, best_complexity)

wclass_tree_final_fit <- fit(wclass_tree_final, data = white_train)

wDecisionTree <- wclass_tree_final_fit %>%
  extract_fit_engine() %>%
  rpart.plot()

# augmented on training 
wDecisionTreeAcc <- augment(wclass_tree_final_fit, new_data = white_train) %>%
  accuracy(truth = quality, estimate = .pred_class)

wDecisionTreeConfMatrix <- augment(wclass_tree_final_fit, new_data = white_train) %>%
  conf_mat(truth = quality, estimate = .pred_class)

# augmented on testing 
wDecisionTreeAccTest <-augment(wclass_tree_final_fit, new_data = white_test) %>%
  conf_mat(truth = quality, estimate = .pred_class) 

wDecisionTreeConfMatrixTest <-augment(wclass_tree_final_fit, new_data = white_test) %>%
  accuracy(truth = quality, estimate = .pred_class)
```

```{r decision tree white LOAD}
# JUST LOAD IN INFO WITH THIS RATHER THAN RERUNNING 
load(file = "WhiteWineDecisionTree.rda")
```

```{r decision tree red}
# decision tree specification
rtree_spec <- decision_tree() %>%
  set_engine("rpart")

# setting mode to classificiation
rtree_spec_class <- rtree_spec %>%
  set_mode("classification")

rclass_tree_fit <- rtree_spec_class %>%
  fit(quality ~ volatile.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = red_train)


rclass_tree_fit %>%
  extract_fit_engine() %>%
  rpart.plot()

# augmentsed on training 
augment(rclass_tree_fit, new_data = red_train) %>%
  accuracy(truth = quality, estimate = .pred_class)

augment(rclass_tree_fit, new_data = red_train) %>%
  conf_mat(truth = quality, estimate = .pred_class)

# augmentsed on testing 
augment(rclass_tree_fit, new_data = red_train) %>%
  conf_mat(truth = quality, estimate = .pred_class) 

augment(rclass_tree_fit, new_data = red_train) %>%
  accuracy(truth = quality, estimate = .pred_class)

# tuning cost complexity 
rclass_tree_wf<- workflow() %>%
  add_model(rtree_spec_class %>% 
              set_args(cost_complexity = tune())) %>% 
  add_formula(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol)

param_grid <- grid_regular(cost_complexity(range = c(-3,-1)), levels = 10)

tune_res_red <- tune_grid(
  rclass_tree_wf,
  resamples = red_fold,
  grid = param_grid,
  metric = metric_set(accuracy)
)

autoplot(tune_res_red)

# extracting the best cost complexitiy parameter

rbest_rocauc <-collect_metrics(tune_res_red) %>% arrange(desc(mean))

best_complexity <- select_best(tune_res_red)

rclass_tree_final <- finalize_workflow(rclass_tree_wf, best_complexity)

rclass_tree_final_fit <- fit(rclass_tree_final, data = red_train)

rclass_tree_final_fit %>%
  extract_fit_engine() %>%
  rpart.plot()

# augmented on training 
augment(rclass_tree_final_fit, new_data = red_train) %>%
  accuracy(truth = quality, estimate = .pred_class)

augment(rclass_tree_final_fit, new_data = red_train) %>%
  conf_mat(truth = quality, estimate = .pred_class)

# augmented on testing 
augment(rclass_tree_final_fit, new_data = red_test) %>%
  conf_mat(truth = quality, estimate = .pred_class) 

augment(rclass_tree_final_fit, new_data = red_test) %>%
  accuracy(truth = quality, estimate = .pred_class)
```

```{r decision tree red LOAD}
# JUST LOAD IN INFO WITH THIS RATHER THAN RERUNNING 
load(file = "RedWineDecisionTree.rda")
```

```{r random forest white }
# setting random forest model up
wrandfor <- rand_forest() %>% set_engine("ranger", importance = "impurity") %>% set_mode("classification")

wrandfor_wf <- workflow() %>%
  add_model(wrandfor %>%
              set_args(mtry = tune(), trees = tune(), min_n = tune())) %>% 
  add_formula(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol)

# tuning the model to find the best arguments 
param_grid1 <- grid_regular(mtry(range = c(1,9)), trees(range = c(15,17)), min_n(range = c(30,50)), levels = 8)

wtune_res_randfor <- tune_grid(
  wrandfor_wf,
  resamples = white_fold,
  grid = param_grid1,
  metric = metric_set(accuracy)
)

wAutoPlotRF <- autoplot(wtune_res_randfor)


# collecting metrics to find best mean
wbest_rocauc1 <-collect_metrics(wtune_res_randfor) %>% arrange(desc(mean))

best_metric <- select_best(wtune_res_randfor)

wrandfor_final <- rand_forest(mtry = 2, trees = 17, min_n = 30) %>% set_engine("ranger", importance = "impurity") %>% set_mode("classification")

wrandfor_fit_final <- fit(wrandfor_final, formula = quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = white_train)

wVIP <- vip(wrandfor_fit_final)
```

```{r random forest white LOAD}
# JUST LOAD IN INFO WITH THIS RATHER THAN RERUNNING 
load(file = "WhiteWineRandomForest.rda")
```

```{r random forest red}
# setting random forest model up
rrandfor <- rand_forest() %>% set_engine("ranger", importance = "impurity") %>% set_mode("classification")

rrandfor_wf <- workflow() %>%
  add_model(rrandfor %>%
              set_args(mtry = tune(), trees = tune(), min_n = tune())) %>% 
  add_formula(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol)

# tuning the model to find the best arguments 
param_grid2 <- grid_regular(mtry(range = c(1,9)), trees(range = c(15,17)), min_n(range = c(30,50)), levels = 8)

rtune_res_randfor <- tune_grid(
  rrandfor_wf,
  resamples = red_fold,
  grid = param_grid2
)

rAutoPlotRF <- autoplot(rtune_res_randfor)

# collecting metrics to find best mean
rbest_rocauc1 <- collect_metrics(rtune_res_randfor) %>% arrange(desc(mean))
print(rbest_rocauc1)
best_metric <- select_best(rtune_res_randfor)

rrandfor_final <- rand_forest(mtry = 7, trees = 17, min_n = 32) %>% set_engine("ranger", importance = "impurity") %>% set_mode("classification")
rrandfor_fit_final <- fit(rrandfor_final, formula = quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol,data = red_train)

rVIP <- vip(rrandfor_fit_final)
```

```{r random forest red LOAD}
# JUST LOAD IN INFO WITH THIS RATHER THAN RERUNNING 
load(file = "RedWineRandomForest.rda")
```

```{r boosted tree white}
wboost_spec <-  boost_tree(tree_depth = 5) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

wboost_wf <- workflow() %>%
  add_model(wboost_spec %>%
              set_args(trees = tune())) %>% 
  add_formula(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol)

param_grid3 <- grid_regular(trees(range = c(10,2000)), levels = 10)

wtune_res_boosted <- tune_grid(
  wboost_wf,
  resamples = white_fold,
  grid = param_grid3
)
autoplot(wtune_res_boosted)
wbest_rocauc2 <- collect_metrics(wtune_res_boosted) %>% arrange(desc(mean))
print(best_rocauc2)
```

```{r boosted tree white LOAD}
# JUST LOAD IN INFO WITH THIS RATHER THAN RERUNNING 
load(file = "WhiteWineBoostedTrees.rda")
```

```{r boosted tree red}
rboost_spec <-  boost_tree(tree_depth = 5) %>% 
  set_engine("xgboost") %>% 
  set_mode("classification")

rboost_wf <- workflow() %>%
  add_model(rboost_spec %>%
              set_args(trees = tune())) %>% 
  add_formula(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol)

param_grid4 <- grid_regular(trees(range = c(10,2000)), levels = 10)

rtune_res_boosted <- tune_grid(
  rboost_wf,
  resamples = red_fold,
  grid = param_grid4
)
autoplot(rtune_res_boosted)

wbest_rocauc2 <- collect_metrics(rtune_res_boosted) %>% arrange(desc(mean))
print(wbest_rocauc2)
```

```{r random forest red LOAD}
# JUST LOAD IN INFO WITH THIS RATHER THAN RERUNNING 
load(file = "RedWineBoostedTrees.rda")
```

```{r best roc}
# augment all the rocs from decision, randforest, boosted --> see which one has the best --> fit best on testing 
# WHITE
wbest_rocauc
wbest_rocauc1
wbest_rocauc2
wbest_roc_table <- rbind(wbest_rocauc[1,c(2,4)], wbest_rocauc1[1,c(4,6)], wbest_rocauc2[1,c(2,4)] ) %>% mutate(model_type = c("Decision Tree", "Random Forest", "Boosted Trees"))
wbest_roc_table

# We will test our model using Boosted Trees and Random Forest 

#RED
rbest_rocauc
rbest_rocauc1
rbest_rocauc2
rbest_roc_table <- rbind(rbest_rocauc[1,c(2,4)], rbest_rocauc1[1,c(4,6)], rbest_rocauc2[1,c(2,4)] ) %>% mutate(model_type = c("Decision Tree", "Random Forest", "Boosted Trees"))
rbest_roc_table

# We will test our model using Boosted Trees and Random Forest 

```

```{r pca with lda white}
white_recipe_pca <- recipe(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = white_train) %>% step_dummy(all_nominal_predictors()) %>% step_normalize(all_predictors()) %>% step_pca(all_numeric_predictors(), num_comp = tune())

# column name(s) must match tune() above
tuneGrid <- expand.grid(num_comp = 1:ncol(white_recipe_pca$template))

# control tune_grid() process below
trControl <- control_grid(verbose = TRUE, allow_par = FALSE)

wlda_pca_wkflow <- workflow() %>% 
  add_model(wlda_model) %>% 
  add_recipe(white_recipe_pca)

pca_lda_fit <- wlda_pca_wkflow %>%
  tune_grid(resamples = white_fold,
            grid = tuneGrid,
            metrics = metric_set(accuracy),
            control = trControl)

pca_lda_metrics <- pca_lda_fit %>% collect_metrics()

ggplot(pca_lda_metrics, aes(x = num_comp, y = mean)) +
  geom_line(color = "#3E4A89FF", linewidth = 2, alpha = 0.6) +
  scale_x_continuous(breaks = 1:ncol(white_recipe_pca$template)) +
  facet_wrap(~.metric) +
  theme_bw()

pca_lda_fit %>% show_best(metric = "accuracy")

(bestTune <- pca_lda_fit %>% select_by_one_std_err(num_comp, metric = "accuracy"))

wlda_pca_wkflow_final <- wlda_pca_wkflow %>% finalize_workflow(bestTune)

fit_final <- wlda_pca_wkflow_final %>% fit(white_train)

white.PCALDA <- tibble(white_train,
       predict(fit_final, new_data = white_train, type = "class"), # predicted class
       predict(fit_final, new_data = white_train, type = "prob"), # posterior prob. for classes
       as_tibble(predict(fit_final, new_data = white_train, type = "raw")$x)) # LD scores
white.PCALDA

# plot
ggplot(white.PCALDA, aes(x = LD1, y = LD2)) +
  geom_point(aes(color = quality, shape = .pred_class)) + 
  theme_bw() +
  ggtitle("PCA-LDA (DAPC) on White Wine Training dataset, using 9 PC")

#augmented on training 
pcalda_fit <- augment(fit_final, new_data = white_train) 
pcalda_acc <- pcalda_fit %>% accuracy(truth = quality, estimate = .pred_class)
pcalda_rocauc <- pcalda_fit %>% roc_auc(truth = quality, estimate = .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% mutate(model_type = "White Wine LDA Model using PCA ")
pcalda_roccurve <- pcalda_fit %>% roc_curve(quality, .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% autoplot()
pcalda_confusionmatrix <- pcalda_fit %>%
  conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

pcalda_acc
pcalda_rocauc
pcalda_roccurve
pcalda_confusionmatrix

#augmented on testing
pcalda_fit_test <- augment(fit_final, new_data = white_test) 
pcalda_test_acc <- pcalda_fit_test %>% accuracy(truth = quality, estimate = .pred_class)
pcalda_test_rocauc <- pcalda_fit_test %>% roc_auc(truth = quality, estimate = .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% mutate(model_type = "White Wine LDA Model using PCA")
pcalda_test_roccurve <- pcalda_fit_test %>% roc_curve(quality, .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% autoplot()
pcalda_test_confusionmatrix <- pcalda_fit_test %>%
  conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

pcalda_test_acc
pcalda_test_rocauc
pcalda_test_roccurve
pcalda_test_confusionmatrix

save(white_recipe_pca,tuneGrid, trControl, wlda_pca_wkflow, pca_lda_fit, pca_lda_metrics, pca_lda_fit, white.PCALDA, pcalda_fit, pcalda_acc, pcalda_rocauc, pcalda_roccurve, pcalda_confusionmatrix, pcalda_test_acc, pcalda_test_rocauc, pcalda_test_roccurve, pcalda_test_confusionmatrix, file = "WhiteLDAPCA.rda")
```

```{r pca with lda red}
red_recipe_pca <- recipe(quality ~ volatile.acidity + fixed.acidity + citric.acid + residual.sugar + chlorides + free.sulfur.dioxide + total.sulfur.dioxide + density + pH + sulphates + alcohol, data = red_train) %>% step_dummy(all_nominal_predictors()) %>% step_normalize(all_predictors()) %>% step_pca(all_numeric_predictors(), num_comp = tune())

# column name(s) must match tune() above
rtuneGrid <- expand.grid(num_comp = 1:ncol(red_recipe_pca$template))

# control tune_grid() process below
rtrControl <- control_grid(verbose = TRUE, allow_par = FALSE)

rlda_pca_wkflow <- workflow() %>% 
  add_model(rlda_model) %>% 
  add_recipe(red_recipe_pca)

rpca_lda_fit <- rlda_pca_wkflow %>%
  tune_grid(resamples = red_fold,
            grid = tuneGrid,
            metrics = metric_set(accuracy),
            control = trControl)

rpca_lda_metrics <- rpca_lda_fit %>% collect_metrics()

ggplot(rpca_lda_metrics, aes(x = num_comp, y = mean)) +
  geom_line(color = "#3E4A89FF", linewidth = 2, alpha = 0.6) +
  scale_x_continuous(breaks = 1:ncol(red_recipe_pca$template)) +
  facet_wrap(~.metric) +
  theme_bw()

rpca_lda_fit %>% show_best(metric = "accuracy")

(bestTune <- rpca_lda_fit %>% select_by_one_std_err(num_comp, metric = "accuracy"))

rlda_pca_wkflow_final <- rlda_pca_wkflow %>% finalize_workflow(bestTune)

rfit_final <- rlda_pca_wkflow_final %>% fit(red_train)

red.PCALDA <- tibble(red_train,
       predict(rfit_final, new_data =red_train, type = "class"), # predicted class
       predict(rfit_final, new_data = red_train, type = "prob"), # posterior prob. for classes
       as_tibble(predict(rfit_final, new_data = red_train, type = "raw")$x)) # LD scores

# plot
ggplot(red.PCALDA, aes(x = LD1, y = LD2)) +
  geom_point(aes(color = quality, shape = .pred_class)) + 
  theme_bw() +
  ggtitle("PCA-LDA (DAPC) on Red Wine Training dataset, using 9 PC")

#augmented on training 
rpcalda_fit <- augment(rfit_final, new_data = red_train) 
rpcalda_acc <- rpcalda_fit %>% accuracy(truth = quality, estimate = .pred_class)
rpcalda_rocauc <- rpcalda_fit %>% roc_auc(truth = quality, estimate = .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% mutate(model_type = "White Wine LDA Model using PCA ")
rpcalda_roccurve <- rpcalda_fit %>% roc_curve(quality, .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% autoplot()
rpcalda_confusionmatrix <- rpcalda_fit %>%
  conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

rpcalda_acc
rpcalda_rocauc
rpcalda_roccurve
rpcalda_confusionmatrix

#augmented on testing
rpcalda_fit_test <- augment(rfit_final, new_data = red_test) 
rpcalda_test_acc <- rpcalda_fit_test %>% accuracy(truth = quality, estimate = .pred_class)
rpcalda_test_rocauc <- rpcalda_fit_test %>% roc_auc(truth = quality, estimate = .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% mutate(model_type = "Red Wine LDA Model using PCA")
rpcalda_test_roccurve <- rpcalda_fit_test %>% roc_curve(quality, .pred_3,.pred_4, .pred_5 , .pred_6 , .pred_7, .pred_8) %>% autoplot()
rpcalda_test_confusionmatrix <- rpcalda_fit_test %>%
  conf_mat(truth = quality, estimate = .pred_class) %>% autoplot(type = "heatmap")

rpcalda_test_acc
rpcalda_test_rocauc
rpcalda_test_roccurve
rpcalda_test_confusionmatrix

save(red_recipe_pca, rtuneGrid, rtrControl, rlda_pca_wkflow, rpca_lda_fit, rpca_lda_metrics, rpca_lda_fit, red.PCALDA, rpcalda_fit, rpcalda_acc, rpcalda_rocauc, rpcalda_roccurve, rpcalda_confusionmatrix, rpcalda_test_acc, rpcalda_test_rocauc, rpcalda_test_roccurve, rpcalda_test_confusionmatrix, file = "RedLDAPCA.rda")
```
